Formal Verification Tool Based on Symbolic Execution for Smart Contract

### 2023-11-28 papers/static-checker/ecoop2024/ecoop2024.tex
add sections
revise the tex file
Add system model

\section{Symblic Execution Model}
\label{sec:symblic-execution-model}

Let $\STACK = (\TermOne, \TYF) \STACKCONCAT (\TermTwo, \TYS) \STACKCONCAT \DOT
\STACKCONCAT \EMPTYSTACK$ be a stack, where elements are types paired with terms.
\\
Let $\INSTRUCTION = \InstructionOne; \InstructionTwo; \DOT; \InstructionN$ be a sequence of instructions. 
\\
Let \PREDICATE\ be a predicate.

\begin{definition}
A system state of the symbolic execution is a tuple \\ $\STATE = [\INSTRUCTION, \STACK, \PREDICATE]$. Let $\SYSTEM = \{\STATEONE, \STATETWO, \DOT, \STATEN \}$ range over sets of system states.
\end{definition}

Let $S_{init}$   = (\KPAIR\ \VPAR\ \VSTORAGE, \TPAIR\ \TYF\ \TYS) \STACKCONCAT\EMPTYSTACK\ be the initial  stack, where \VPAR\ and \VSTORAGE\ are the terms that represent the parameter and the storage and \TYF\ and \TYS\ are their corresponding types, and  $S_{final}$   = (\PAIR\ \VOPERATIONLIST\ \VSTORAGE, \TPAIR\ (\TOPERATIONLIST) \TYS) \STACKCONCAT\EMPTYSTACK\ be  the final stack, where \VOPERATIONLIST\ repesents a operation list

\begin{figure}[h]
\begin{align*}
\text{p} &::= \\
   &\Mid\ \langle \text{atomic formula} \rangle \\
   &\Mid\ \NEG\ \text{p} \
   \Mid\ \text{p}\ \CNF\ \text{q} \
   \Mid\ \text{p}\ \DNF\ \text{q} \\
\langle \text{atomic formula} \rangle &::= \\ 
    &\Mid\ \langle \text{butop} \rangle \langle \text{bterm} \rangle\\ 
   &\Mid\  \langle \text{bterm} \rangle  \langle \text{biop} \rangle \langle \text{bterm} \rangle\\ 
\langle \text{butop} \rangle &::= \\
   &\Mid\ \text{not} \\
\langle \text{biop} \rangle &::= \\
   &\Mid\ \text{$=$} \
   \Mid\ \text{$>$} \
   \Mid\ \text{$<$} \
   \Mid\ \text{$>=$} \
   \Mid\ \text{$=<$} \
   \Mid\ \text{$!=$} \\
   &\Mid\ \text{and}\ \Mid\ \text{or}\ \Mid\ \text{xor}\\
\langle \text{bterm} \rangle &::= \\ 
   &\Mid\ \text{t} \\
   &\Mid\ \langle \text{unop} \rangle \text{t} \\ 
   &\Mid\ \text{t}\ \langle \text{binop} \rangle \text{t}\\  
\langle \text{unop} \rangle &::= \\ 
   &\Mid\ \text{abs} \\
   &\Mid\ \text{size} \\
   &\Mid\ \text{int} \\
   &\Mid\ \text{contract}  \langle \text{type} \rangle \\
   &\Mid\ \text{isleft} \\
   &\Mid\ \text{isnone} \\
   &\Mid\ \text{neg} \\
   &\Mid\ \text{blake2b} \\
   &\Mid\ \text{hash-key} \\
   &\Mid\ \text{keccak} \\
   &\Mid\ \text{pairing-check} \\
   &\Mid\ \text{sha256} \\
   &\Mid\ \text{sha3} \\
   &\Mid\ \text{sha512} \\
   &\Mid\ \text{implicit-account} \\
   &\Mid\ \text{check-sig} \\
\langle \text{bitop} \rangle &::= \\ 
   &\Mid\ \text{+}\ \Mid\ \text{-}\ \Mid\ \text{*}\ \Mid\ \text{/}
\end{align*}
\caption{Predicates}
\label{fig:predicate}
\end{figure}

@@

\pagebreak
\subsection{Rules}
The rule semantic is defined by several kinds of transitions:
\begin{enumerate}
\item \ExprTrans\ single-step evaluation of an expression in a system state,
\item \StateTrans\ internal transitions of a system state, which is non-deterministic  as shown for instructions \ITER\ and \LOOP,
\item \SystemTrans\ symbolic system transitions.
\end{enumerate}

\subsubsection{System rules}
\begin{mathpar}
\inferrule[INVALID-PRE]
  { \NEG\ \PREDICATE
  }{
  \{[\INSTRUCTION, \STACK, \PREDICATE]\} \cup \SYSTEM \SystemTrans \SYSTEM}
\end{mathpar}

\subsubsection{Instruction rules}
\paragraph{Control structures}
%EXE
\begin{mathpar}
  \inferrule[EXEC]
  {
    [\INSTRUCTIONONE, (\StackOne, \TYF) \STACKCONCAT \EMPTYSTACK, 
    Q]
    \StateTrans^*
    [ \EMPTYSTACK, (\StackOne', \TYS) \STACKCONCAT Q']
  }{
     [(\EXEC; \INSTRUCTION),   (\{\INSTRUCTIONONE\}, \TYF\ \rightarrow\ \TYS) \STACKCONCAT
    (\StackOne, \TYF) \STACKCONCAT \STACK, \PREDICATE \wedge Q] 
    \StateTrans \
    [ \INSTRUCTION, (\StackOne', \TYS) \STACKCONCAT \STACK,
    \PREDICATE \wedge Q']}
\end{mathpar}

@@

%IF
\begin{mathpar}
  \inferrule[IF-true]
  {  
  }{
    [(\IF\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TBOOL) \STACKCONCAT\STACK, \PREDICATE]
    \StateTrans\
    [\INSTRUCTIONONE, \STACK, \PREDICATE\ \Wedge\ \StackOne]
  }

%LOOP
\begin{mathpar}
  \inferrule[LOOP-true]
  {  
  }{
    [(\LOOP\ \INSTRUCTIONONE; \INSTRUCTION),  (\StackOne, \TBOOL)
    \STACKCONCAT\STACK, \PREDICATE]
    \StateTrans\
    [(\INSTRUCTIONONE; \LOOP\ \INSTRUCTIONONE; \INSTRUCTION),
    \STACK, \PREDICATE \wedge \StackOne]
  }

  \inferrule[LOOP-false]
  {  
  }{
    [(\LOOP\ \INSTRUCTIONONE; \INSTRUCTION),  (\StackOne, \TBOOL) \STACKCONCAT\
    \STACK, \PREDICATE]
    \StateTrans\
   [\INSTRUCTION, \STACK, \PREDICATE \wedge
   (\NEG\StackOne)]
   }
\end{mathpar}

%%ITER
\begin{mathpar}
  \inferrule[ITER-empty]
  {  
  }{
    [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE] \StateTrans\ [\INSTRUCTION, \STACK,  \PREDICATE\ \Wedge\  (\StackOne\ \EQUAL\ \EMPTYLIST)]
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[ITER-nonempty]
  {  [\ITER,  (\HEAD, \TY) \STACKCONCAT\STACK, 
    Q]
    \StateTrans^*
    [ \EMPTYSTACK,  \STACK', Q']
  }{
    [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE \wedge Q] \StateTrans \\ [(\TITER\ \INSTRUCTIONONE ; \INSTRUCTION), \{\STAIL\}\ \STACKCONCAT\STACK',  \PREDICATE\ \Wedge  ( \StackOne\ \EQUAL\ \{\HEAD; \STAIL \})  \wedge Q']
  }
\end{mathpar}

@@

@@

\subsubsection{Operation on tickets}
\subsubsection{FAILWITH}
%FAILWITH
\begin{mathpar}
  \inferrule[FAILWITH]
  {
  }{[(\FAILWITH; \INSTRUCTION), (\StackOne,  \TY) \STACKCONCAT \STACK,  \PREDICATE] \StateTrans [\EMPTYSTACK, failwith (\StackOne) \STACKCONCAT\EMPTYSTACK, \PREDICATE]}
\end{mathpar}
\section {Domain Specific Language}
\section {Static Checker}
\section {Case Studies}
\section {Related Work}
\section {Conclusion}
\end{document}

>>>>

\section{Symbolic Execution Model}
\label{sec:symbolic-execution-model}
\subsection{System Model}
Michelson is the low-level smart contract language employed in the Tezos blockchain platform. It is characterized as a stack-based language, featuring high-level data types and primitives and integrates strict static type checking into its stack-based execution mode.

The processing and manipulation of data during the execution of smart contracts are facilitated using a stack. A stack  can be written:
\begin{itemize}
\item[]  \EMPTYSTACK\ for the empty stack.
\item[]  \HEAD\ \STACKCONCAT\ \NTAIL\ for the stack whose first value has type (\HEAD) and queue has stack type (\NTAIL).
\end{itemize}

Each piece of data in Michelson is assigned a specific type, and the language imposes restrictions preventing programmers from introducing custom types. Consider a stack denoted as $\STACK = (\TermOne, \TYF) \STACKCONCAT (\TermTwo, \TYS) \STACKCONCAT \DOT
\STACKCONCAT \EMPTYSTACK$ where elements are types paired with terms.

A Michelson program is a sequence of instructions executed in a sequential manner. Each instruction takes, as input, the stack produced by the preceding instruction and modifies it for the subsequent one. This sequential flow ensures that each instruction builds upon the state left by its predecessor in the program. Let $\INSTRUCTION = \InstructionOne; \InstructionTwo; \DOT; \InstructionN$ represent a sequence of instructions.

While undergoing symbolic execution, branching may occur depending on the conditions, and these branching conditions are recorded as predicates. Let \PREDICATE\ represent a predicate expressed in conjunction form.

\begin{definition}
A symbolic execution's state is represented by the tuple $\STATE = [\INSTRUCTION, \STACK, \PREDICATE]$. Here, $\SYSTEM = \{\STATEONE, \STATETWO, \DOT, \STATEN \}$ range over sets of  states.
\end{definition}

The Michelson program takes as input a stack that consists of a single pair whose first element is an input value referred to as the parameter, and second element represents the content of the storage space. The program produces a stack as output, containing a single pair. The first element of this pair is the list of internal operations that the program intends to emit, and the second element is the new contents of the storage space.

Define the initial stack as $S_{init}$   = (\KPAIR\ \VPAR\ \VSTORAGE, \TPAIR\ \TYF\ \TYS) \STACKCONCAT\EMPTYSTACK, where \VPAR\ and \VSTORAGE\ denote terms representing the parameter and the storage, and \TYF\ and \TYS\ are their corresponding types. Similarly, designate the final stack as $S_{final}$   = (\PAIR\ \VOPERATIONLIST\ \VSTORAGE, \TPAIR\ (\TOPERATIONLIST) \TYS) \STACKCONCAT\EMPTYSTACK, where \VOPERATIONLIST\ repesents a operation list.

@@

\pagebreak
\subsection{Rules}

Each instruction is defined by a rule on the system model. The rule semantic is defined by several kinds of transitions:
\begin{enumerate}
\item \ExprTrans\ single-step evaluation of an expression in a system state,
\item \StateTrans\ internal transitions of a state, which is non-deterministic  as shown for loop instructions, such as \LOOP\ and \LOOP-LEFT,
\item \SystemTrans\ symbolic system transitions.
\end{enumerate}

During the execution of symbolic analysis, numberous states are generated. However, certain states may be unreachable from a symbolic initial state. This occurs when the predicate \PREDICATE\ is unsatisfied. Consequently, a system state is disregarded when the predicate \PREDICATE\ is fail.

\begin{mathpar}
\inferrule[]
  { \NEG\ \PREDICATE
  }{
  \{[\INSTRUCTION, \STACK, \PREDICATE]\} \cup \SYSTEM \SystemTrans \SYSTEM}
\end{mathpar}
There are various types of instructions in Michelson, categorized based on their functions. These include control structures, stack manipulation, arithmetic operations, boolean operations, cryptographic operations, and operations on data structures. In our model, we classify them into three main groups: basic instructions, branch instructions, and loop instructions. This categorization reflects how these instructions are handled within our specific model.
\subsubsection{Basic Instructions}

\paragraph{Control structures}
%EXE
\begin{mathpar}
  \inferrule[EXEC]
  {
    [\INSTRUCTIONONE, (\StackOne, \TYF) \STACKCONCAT \EMPTYSTACK, 
    Q]
    \StateTrans^*
    [ \EMPTYSTACK, (\StackOne', \TYS) \STACKCONCAT Q']
  }{
     [(\EXEC; \INSTRUCTION),   (\{\INSTRUCTIONONE\}, \TYF\ \rightarrow\ \TYS) \STACKCONCAT
    (\StackOne, \TYF) \STACKCONCAT \STACK, \PREDICATE \wedge Q] 
    \StateTrans \
    [ \INSTRUCTION, (\StackOne', \TYS) \STACKCONCAT \STACK,
    \PREDICATE \wedge Q']}
\end{mathpar}

@@

@@

\subsubsection{Branch Instructions}
%IF
\begin{mathpar}
  \inferrule[IF-true]
  {  
  }{
    [(\IF\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TBOOL) \STACKCONCAT\STACK, \PREDICATE]
    \StateTrans\
    [\INSTRUCTIONONE, \STACK, \PREDICATE\ \Wedge\ \StackOne]
  }

\subsubsection{Loop Instructions}
\paragraph{While loops}
%LOOP
\begin{mathpar}
  \inferrule[LOOP-true]
  {  
  }{
    [(\LOOP\ \INSTRUCTIONONE; \INSTRUCTION),  (\StackOne, \TBOOL)
    \STACKCONCAT\STACK, \PREDICATE]
    \StateTrans\
    [(\INSTRUCTIONONE; \LOOP\ \INSTRUCTIONONE; \INSTRUCTION),
    \STACK, \PREDICATE \wedge \StackOne]
  }

  \inferrule[LOOP-false]
  {  
  }{
    [(\LOOP\ \INSTRUCTIONONE; \INSTRUCTION),  (\StackOne, \TBOOL) \STACKCONCAT\
    \STACK, \PREDICATE]
    \StateTrans\
   [\INSTRUCTION, \STACK, \PREDICATE \wedge
   (\NEG\StackOne)]
   }
\end{mathpar}
\paragraph{For loops}
%%ITER
\begin{mathpar}
  \inferrule[ITER-empty]
  {  
  }{
    [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE] \StateTrans\ [\INSTRUCTION, \STACK,  \PREDICATE\ \Wedge\  (\StackOne\ \EQUAL\ \EMPTYLIST)]
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[ITER-nonempty]
  {  [\ITER,  (\HEAD, \TY) \STACKCONCAT\STACK, 
    Q]
    \StateTrans^*
    [ \EMPTYSTACK,  \STACK', Q']
  }{
    [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE \wedge Q] \StateTrans \\ [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), \{\STAIL\}\ \STACKCONCAT\STACK',  \PREDICATE \wedge Q' \Wedge  ( \StackOne\ \EQUAL\ \{\HEAD; \STAIL \}) ]
  }
\end{mathpar}

@@

\subsubsection{Operation on tickets}
\subsubsection{FAILWITH}
a Michelson program can fail, explicitly using a specific opcode

%FAILWITH
\begin{mathpar}
  \inferrule[FAILWITH]
  {
  }{[(\FAILWITH; \INSTRUCTION), (\StackOne,  \TY) \STACKCONCAT \STACK,  \PREDICATE] \StateTrans [\EMPTYSTACK, failwith (\StackOne) \STACKCONCAT\EMPTYSTACK, \PREDICATE]}
\end{mathpar}
\subsection{Dealing with loops}
\subsubsection{For loop: ITER}
\subsubsection{While loops: \LOOP\ and \LOOP-LEFT}
\section {Domain Specific Language}
\section {Static Checker}
\section {Case Studies}
\subsection{USDtz}
\subsection{Kolibri Oracle Contract}
\section {Related Work}
\section {Conclusion}
\section {Appendix}

\begin{figure}[h]
\begin{align*}
\text{p} &::= \\
   &\Mid\ \langle \text{atomic formula} \rangle \\
   &\Mid\ \NEG\ \text{p} \
   \Mid\ \text{p}\ \CNF\ \text{q} \
   \Mid\ \text{p}\ \DNF\ \text{q} \\
\langle \text{atomic formula} \rangle &::= \\ 
    &\Mid\ \langle \text{butop} \rangle \langle \text{bterm} \rangle\\ 
   &\Mid\  \langle \text{bterm} \rangle  \langle \text{biop} \rangle \langle \text{bterm} \rangle\\ 
\langle \text{butop} \rangle &::= \\
   &\Mid\ \text{not} \\
\langle \text{biop} \rangle &::= \\
   &\Mid\ \text{$=$} \
   \Mid\ \text{$>$} \
   \Mid\ \text{$<$} \
   \Mid\ \text{$>=$} \
   \Mid\ \text{$=<$} \
   \Mid\ \text{$!=$} \\
   &\Mid\ \text{and}\ \Mid\ \text{or}\ \Mid\ \text{xor}\\
\langle \text{bterm} \rangle &::= \\ 
   &\Mid\ \text{t} \\
   &\Mid\ \langle \text{unop} \rangle \text{t} \\ 
   &\Mid\ \text{t}\ \langle \text{binop} \rangle \text{t}\\  
\langle \text{unop} \rangle &::= \\ 
   &\Mid\ \text{abs} \\
   &\Mid\ \text{size} \\
   &\Mid\ \text{int} \\
   &\Mid\ \text{contract}  \langle \text{type} \rangle \\
   &\Mid\ \text{isleft} \\
   &\Mid\ \text{isnone} \\
   &\Mid\ \text{neg} \\
   &\Mid\ \text{blake2b} \\
   &\Mid\ \text{hash-key} \\
   &\Mid\ \text{keccak} \\
   &\Mid\ \text{pairing-check} \\
   &\Mid\ \text{sha256} \\
   &\Mid\ \text{sha3} \\
   &\Mid\ \text{sha512} \\
   &\Mid\ \text{implicit-account} \\
   &\Mid\ \text{check-sig} \\
\langle \text{bitop} \rangle &::= \\ 
   &\Mid\ \text{+}\ \Mid\ \text{-}\ \Mid\ \text{*}\ \Mid\ \text{/}
\end{align*}
\caption{Predicates}
\label{fig:predicate}
\end{figure}
\subsubsection{Basic Instructions}

\paragraph{Control structures}
%EXE
\begin{mathpar}
  \inferrule[EXEC]
  {
    [\INSTRUCTIONONE, (\StackOne, \TYF) \STACKCONCAT \EMPTYSTACK, 
    Q]
    \StateTrans^*
    [ \EMPTYSTACK, (\StackOne', \TYS) \STACKCONCAT Q']
  }{
     [(\EXEC; \INSTRUCTION),   (\{\INSTRUCTIONONE\}, \TYF\ \rightarrow\ \TYS) \STACKCONCAT
    (\StackOne, \TYF) \STACKCONCAT \STACK, \PREDICATE \wedge Q] 
    \StateTrans \
    [ \INSTRUCTION, (\StackOne', \TYS) \STACKCONCAT \STACK,
    \PREDICATE \wedge Q']}
\end{mathpar}

\subsubsection{Branch Instructions}
%IF
\begin{mathpar}
  \inferrule[IF-true]
  {  
  }{
    [(\IF\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TBOOL) \STACKCONCAT\STACK, \PREDICATE]
    \StateTrans\
    [\INSTRUCTIONONE, \STACK, \PREDICATE\ \Wedge\ \StackOne]
  }

\subsubsection{Loop Instructions}
\paragraph{While loops}
%LOOP
\begin{mathpar}
  \inferrule[LOOP-true]
  {  
  }{
    [(\LOOP\ \INSTRUCTIONONE; \INSTRUCTION),  (\StackOne, \TBOOL)
    \STACKCONCAT\STACK, \PREDICATE]
    \StateTrans\
    [(\INSTRUCTIONONE; \LOOP\ \INSTRUCTIONONE; \INSTRUCTION),
    \STACK, \PREDICATE \wedge \StackOne]
  }

  \inferrule[LOOP-false]
  {  
  }{
    [(\LOOP\ \INSTRUCTIONONE; \INSTRUCTION),  (\StackOne, \TBOOL) \STACKCONCAT\
    \STACK, \PREDICATE]
    \StateTrans\
   [\INSTRUCTION, \STACK, \PREDICATE \wedge
   (\NEG\StackOne)]
   }
\end{mathpar}
\paragraph{For loops}
%%ITER
\begin{mathpar}
  \inferrule[ITER-empty]
  {  
  }{
    [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE] \StateTrans\ [\INSTRUCTION, \STACK,  \PREDICATE\ \Wedge\  (\StackOne\ \EQUAL\ \EMPTYLIST)]
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[ITER-nonempty]
  {  [\ITER,  (\HEAD, \TY) \STACKCONCAT\STACK, 
    Q]
    \StateTrans^*
    [ \EMPTYSTACK,  \STACK', Q']
  }{
    [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE \wedge Q] \StateTrans \\ [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), \{\STAIL\}\ \STACKCONCAT\STACK',  \PREDICATE \wedge Q' \Wedge  ( \StackOne\ \EQUAL\ \{\HEAD; \STAIL \}) ]
  }
\end{mathpar}

\subsubsection{Operation on tickets}
\subsubsection{FAILWITH}
a Michelson program can fail, explicitly using a specific opcode

%FAILWITH
\begin{mathpar}
  \inferrule[FAILWITH]
  {
  }{[(\FAILWITH; \INSTRUCTION), (\StackOne,  \TY) \STACKCONCAT \STACK,  \PREDICATE] \StateTrans [\EMPTYSTACK, failwith (\StackOne) \STACKCONCAT\EMPTYSTACK, \PREDICATE]}
\end{mathpar}
\subsection{Dealing with loops}
\subsubsection{For loop: ITER}
\subsubsection{While loops: \LOOP\ and \LOOP-LEFT}
\end{document}
